---
title: "GIS map using tmap"
author: "Sameer Padhye"
date: "2018-11-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tmap)
require(raster)
require(ggmap)
require(sp)
require(rgdal)
require(rasterVis)
require(viridisLite)
require(RColorBrewer)
```

#Loading the shapefiles or polygons for plotting

The shapefiles should be imported using the *readOGR* function from the *sp* package and provided with an appropriate **CRS** (Coordinate Reference System).

```{r,echo=T,message=FALSE,warning=FALSE}

# India map with states
india_state_map<-readOGR('C:/Data/GIS_layers/IND_adm/IND_adm1.shp')
proj4string(india_state_map)<- CRS("+proj=longlat +datum=WGS84")

# Western Ghats shapefile
Western_Ghats<-readOGR("C:/Data/GIS_layers/WG new outline shapefiles/Western GHats outline.shp")
proj4string(Western_Ghats)<- CRS("+proj=longlat +datum=WGS84")

#Alternatively, we can specify the path to the shapefile separately using the 'dsn' arguement in the readOGR function. The code would be: readOGR(dsn="C:/mapdata",layer='IND_adm1')

```

#Exploring the data

A *Spatialpolygondataframe* object is generated which can then be explored for more details. Note that this object is an S4 object and has *slots* which can be viewed using the **@** operator.

```{r,echo=T,message=FALSE,warning=FALSE}

head(india_state_map@data,
     3)
```

#Extracting specific polygons from the main map

After exploring the object, the specific polygon file to be mapped can be extracted from the main object using the *$* operator. Here, I have extracted Maharashtra and Goa state polygons from the India map

```{r,echo=T}

#Maharashtra state polygon
is_mh_state <- india_state_map$NAME_1== "Maharashtra"
maharashtra_state<-india_state_map[is_mh_state,]

#Goa state polygon
is_goa_state <- india_state_map$NAME_1== "Goa"
goa_state<-india_state_map[is_goa_state,]

```

#Importing GIS data of the samples

Data points to be mapped must also be converted into a *Spatialpolygondataframe* object

```{r,echo=T}
# Importing the data
GIS_samples<-read.csv(file="C:/Data/Research data/Large Branchiopoda/Large branchipod distribution data paper/L_branchiopod data.csv",
                 header=T)
```

The names of the field for the Latitude and Longitude are as per the specific dataset, though, the order should not change. The **CRS** should be the same as used for the map.

```{r,echo=T}

# converting the data into a spatialdataframe object. 
sampling_points<-SpatialPointsDataFrame(coords =GIS_samples[,c("Longitude","Latitude")],                                     data = GIS_samples,
                                        proj4string = CRS("+proj=longlat +datum=WGS84"))
```

#Plotting the map

Plots can then be made as requirement (Here, I have used the *tmap* package, though, maps can also be made using many other libraries like *ggmap*).

```{r,echo=F}

#1. Outline map of states

outline_map<-tm_shape(maharashtra_state)+
  tm_borders()+
  tm_fill(col='grey',alpha = 0.6)+
  tm_shape(Western_Ghats)+
  tm_borders()+
  tm_fill(col='forestgreen')+
  tm_shape(goa_state)+
  tm_borders()+
  tm_fill(col='grey',alpha = 0.6)
  
outline_map


```


```{r,echo=T}

#2. Mapping the sample GIS data on the outline map

sample_map<-outline_map+
  tm_shape(sampling_points)+
  tm_dots(size=0.4, 
          col="red")

sample_map
```

Additional features such as compass, scales and grids can then be added on the map.

```{r,echo=T}

sample_map+
  tm_compass(position = c(.75, .15), 
             color.light = "grey90")+
  tm_layout(inner.margins=c(.11,.05, .04, .04))+ 
  tm_scale_bar(position=c("right", 
                          "bottom"))+
  tm_grid(n.x=4,
          n.y=4,
          lwd=0.4,
          alpha = 0.6,
          col='grey60',
          labels.size = 0.8)

```


Using tmap function 'ttm()' map view can be shifted to an interactive from static and vice versa.

This gives a basic flowchart to generate simple maps. There are many other custom function that can be used to change/modify or beautify the maps. One example is using Raster files along with the shapefiles for providing additional information


##END



