---
title: "C.hislopi_random_points"
author: "Sameer Padhye"
date: "9/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generating random points based on grids
Each locality will be assigned a respective grid after which a set of 200 random spatial points will be generated per grid.Here, grids of 100 sq.km. have been used

Obtaining the grids
```{r,echo=T,message=FALSE,warning=FALSE}

##Grids (100k sq.km)

grids_100k<-readOGR("C:/Users/samee/Downloads/HWSD_RASTER/world_mgrs/mgrs_region.shp")

proj4string(grids_100k)<- CRS("+proj=longlat +datum=WGS84")


#names of the countries and respective grids for convenience

names_grids_w_country<-raster::intersect(c.hislopi_points,grids_100k)

names_grids_w_country@data$GRID1MIL


#extracting the grids specific to the localities

grids_for_points<-raster::intersect(grids_100k,c.hislopi_points)


### REMOVE Palearctic and Nearctic locality from the data

```

Since 8 grids are involved, a function written to extract random spatial points as per requirement

```{r,echo=T,message=FALSE,warning=FALSE}

#Function

random_sp_pts<-function(x,grid_no,s_no){
    
    points=x[grid_no,]%>%
        spsample(.,n=s_no,"random")%>% 
        .@coords%>%
        as.data.frame(.)%>%
        cbind(grid_no=x$GRID1MIL[[grid_no]]%>%
                  rep(.,times=s_no))%>%
        droplevels(.)
    
    return(points)

}

#Using the function to get data for all 8 grids
grid_1<-random_sp_pts(grids_for_points,1,100)
grid_2<-random_sp_pts(grids_for_points,2,100)
grid_3<-random_sp_pts(grids_for_points,3,100)
grid_4<-random_sp_pts(grids_for_points,4,100)
grid_5<-random_sp_pts(grids_for_points,5,100)
grid_6<-random_sp_pts(grids_for_points,6,100)
grid_7<-random_sp_pts(grids_for_points,7,100)
grid_8<-random_sp_pts(grids_for_points,8,100)

```

Obtaining a complete dataset of all grids

```{r,echo=T,message=FALSE,warning=FALSE}

#combining all the 8 grids data

grid_all_data<-as.data.frame(rbind(grid_1,
                                   grid_2,
                                   grid_3,
                                   grid_4,
                                   grid_5,
                                   grid_6,
                                   grid_7,
                                   grid_8))


# selecting the names of the countries and respective grid code from 'names_grids_w_country'

names_with_codes<-names_grids_w_country@data[,c("Country","GRID1MIL")]

#joining the two datasets 

all_data_randompts<-full_join(grid_all_data,
                              names_with_codes, 
                              by = c("grid_no" = "GRID1MIL"))%>%
    dplyr::rename('Longitude'='x',
                  "Latitude"='y')%>%
    drop_na()

head(all_data_randompts,10)

View(all_data_randompts)

```

Obtaining geochem,bioclim,altitude and Ecoregion data for the random points

```{r,echo=T,message=FALSE,warning=FALSE}

# Obtaining a spatialdataframe object of the GIS data of the random points

c.hislopi_points_random<-SpatialPointsDataFrame(coords = all_data_randompts[,c("Longitude","Latitude")],
                                         data = all_data_randompts,
                                         proj4string = CRS("+proj=longlat +datum=WGS84"))


#Geochem data

geochem_data_random<-geochem_files%>%
    purrr::map(raster)%>%
    purrr::map(.,~raster::extract(.,all_data_randompts[,c("Longitude","Latitude")]))%>%
    bind_cols(.)%>%
    purrr::set_names(geochem_colnames)

#Bioclim

random_pts_bioclim<- raster::extract(bioclim,all_data_randompts[,c("Longitude","Latitude")])
 
# Altitude

random_pts_alt<- raster::extract(alt,all_data_randompts[,c("Longitude","Latitude")])


# #Ecoregions
# 
random_pts_data_eco<-over(c.hislopi_points_random, world_ecoreg)

##Combining all data

all_gis_data<-all_data_randompts%>%
    cbind(.,random_pts_bioclim)%>%
    cbind(.,geochem_data_random)%>%
    as.data.frame(.)%>%
    dplyr::mutate(grid_no=all_data_randompts$grid_no,
                  country=all_data_randompts$Country,
                  altitude=random_pts_alt,
                  realm=random_pts_data_eco$WWF_REALM2,
                  ecoregions=random_pts_data_eco$WWF_MHTNAM
                  )%>%
    dplyr::select(grid_no,
                  country,
                  everything())%>%
    #dplyr::select_at(vars(starts_with('therm')))
    na.omit(.)%>%
    droplevels(.)%>%
    dplyr::select(-c("therm_cap0_",
                     "therm_cap10_",
                     "therm_cap100_",
                     "therm_cap50_"))
    

head(all_gis_data,6)
# #recoding realm category
# 
# all_gis_data$realm<-fct_recode(all_gis_data$realm,
#                                Neotropics="Neotropic")


#renaming bioclim names

names(all_gis_data)[6:24]<-c(sprintf("Temp%d",seq=(1:11)),
                                   sprintf("Prec%d",seq=(1:8)))


```


```{r,echo=T,message=FALSE,warning=FALSE}

#Finding collinearity

chislopi_collinearity_random<-all_gis_data%>%
    select_if(is.numeric)%>%
    as.matrix(.)%>%
    cor(.,method = "spearman")

## Removing highly correlated variables using caret package

require(caret)

#Index for selecting high correlated variables

index_selection_random<-findCorrelation(chislopi_collinearity_random,
                                 cutoff = 0.65)%>%
    sort(.)

#Obtaining the variable names after removing the highly correlated variables

final_env_var_chislop_random<-chislopi_collinearity_random[,-c(index_selection_random)]%>%
    data.frame(.)%>%
    mutate_if(is.numeric,
              round, 3)%>%
    colnames(.)

# The v
final_env_var_chislop_random



```


```{r,echo=T,message=FALSE,warning=FALSE}

#selecting the variables having pairwise collinearity less than 0.6

pca_c.hislopi_random <- all_gis_data%>%
    dplyr::select(c(final_env_var_chislop_random))%>%
    prcomp(.,scale. = T)


#extracting the PCA co-ordinates of sites 

pca_coord_chislopi_random<-pca_c.hislopi$x

#extracting PCA vector values of the first two axes

pca.vectors.chislopi_random<-pca_c.hislopi$rotation[,c(1,2)]

#write.csv(pca.vectors.chislopi,'pca_chislopi_vectors.csv')

#plot for PCA

autoplot(pca_c.hislopi_random, 
         scale = 0, 
         data = all_gis_data, 
         colour='realm',
         label=T,
         label.label = "Country",
         size = 5, 
         shape='realm',
         frame=T,
         frame.colour = 'realm',
         loadings = TRUE, 
         loadings.colour = 'black', 
         loadings.label = TRUE, 
         loadings.label.size = 4, 
         loadings.label.hjust = 0.5, 
         loadings.label.vjust = 1.2)+
    theme_bw(base_size = 18)+
    scale_color_manual(values=c("orange",
                                'forestgreen',
                                "blue",
                                "grey50",
                                "black"))+
    scale_size_manual(values =2)+
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank())

#b. Scree plot for PCA

library(factoextra)

fviz_eig(pca_c.hislopi)

```


```{r,echo=T,message=FALSE,warning=FALSE}

# Since Palearctic and Nearctic realm has only one locality, it has been removed for PERMANOVA)

#Getting the data

chislopi_permanova.data<-C.hislopi_all.data%>%
    dplyr::select(realm,
                  c(final_env_var_chislop))%>%
    #mutate_if(is.numeric,log1p)%>%
    na.omit(.)

# checking multivariate spread (condition for permanova) for the cladoceran data

#converting it to a distance object

chislopi_beta<-vegdist(subset(chislopi_permanova.data,select=-c(realm)), 
                       method = "gower",
                       binary = T,
                       na.rm = T)

#Beta dispersion test to check homogeneity of spread

chislopi_betadis<-betadisper(chislopi_beta, 
                             chislopi_permanova.data$realm)

anova(chislopi_betadis)

#permutation test (additional) to test the siginificance

permutest(chislopi_betadis, 
          pairwise = TRUE)

#PERMANOVA

chislopi_permanova<-adonis(subset(chislopi_permanova.data,select=-c(realm)) ~ realm,
                           data = chislopi_permanova.data, 
                           permutations = 5000, 
                           method = "gower")

#PERMANOVA results

chislopi_permanova$aov.tab

```


```{r,echo=T,message=FALSE,warning=FALSE}

```


```{r,echo=T,message=FALSE,warning=FALSE}

```


```{r,echo=T,message=FALSE,warning=FALSE}

```
