---
title: "Leaflet maps"
author: "Sameer Padhye"
date: "5/22/2019"
output: github_document
---
#Using Leaflet package to visualize interactive maps is given below

```{r setup, include=FALSE}
library(leaflet)
library(readxl)
library(tidyverse)
library(rgdal)

```

#Locality data used here is taken from the disk

```{r cars,echo=FALSE}

data_file<-"C:/Users/samee/Desktop/R data/datasets/leaflet_GIS_data.xlsx"

locality_data<-read_excel(data_file,
                            sheet=1)%>%
    dplyr::select(Longitude,
                  Latitude,
                  everything())

```

#Exploring the data

```{r,echo=FALSE}
head(locality_data,5)
```
#Making Leaflet maps using some already available basemaps (*CartoDB.Positron*)
```{r,echo=FALSE}

locality_map1 <- leaflet(locality_data) %>%
    addProviderTiles("CartoDB.Positron") %>%
    addMarkers(lng = ~Longitude, 
               lat = ~Latitude,
               popup = ~Localities)
locality_map1


```

#Making Leaflet maps using some already available basemaps (*Stamen.Terrain*). Colors for the markers are defined earlier

```{r,echo=FALSE}
colour_combo <- colorFactor(c("blue", "green"),
                            domain = unique(locality_data$Country))

locality_map2 <- leaflet(locality_data) %>%
    addProviderTiles("Stamen.Terrain") %>%
    addCircleMarkers(
        color = ~colour_combo(Country),
        opacity = 1,
        stroke = TRUE,
        lng = ~Longitude, 
        lat = ~Latitude,
        label = ~as.character(Localities),
        radius = 4)
locality_map2 

```

#Using an available shapefile on the disk

```{r,echo=FALSE,warning=FALSE}
shapefile_path<-"C:/Users/samee/Downloads/HWSD_RASTER/world_shapefile/ne_50m_admin_0_countries.shp"

world_shape<-readOGR(shapefile_path)

proj4string(world_shape)<- CRS("+proj=longlat +datum=WGS84")

```

#The shapefile is then used for creating the map

```{r,echo=FALSE,warning=FALSE}
world_map <- leaflet(world_shape)

world_map%>%
    addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 0.5,color = grey)%>%
    addCircleMarkers(
        stroke = TRUE,
        lng = ~locality_data$Longitude, 
        lat = ~locality_data$Latitude,
        label = ~as.character(locality_data$Localities),
        radius = 6)
```


#The pointers (markers) can be edited based on shapes and sizes.This part will be added soon
